---
title: "Question 1"
date: "2/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Question 1: Does typical compensation for Data Science professionals differ among densely populated and sparsely populated areas? (‘Dense’ definition >12,000 per square mile; ‘sparse’ definition <6,000 per square mile). **

### *Define density by following metrics country population, number of entries per country, number of entries per title (i.e 7000 Data Scientist vs 3000 Data Engineer)*

### *Use ANOVA to test possible differences in multiple groups and Linear Regression to test if there is a linear relationship between compensation and the density of groups. *

### *Test level of power data set allows for our research given the n number of responses.*

```{r}
# Why did you choose a split of 500 for the density?
data <- read.csv("../Data/Data_Science_Density.csv")
high_density <- data[data$Density > 500,]
low_density <- data[data$Density < 500,]
```

# Interesting fact: We have significantly more datapoints for low density locations than high density locations.

```{r}
# Random data facts (mean, sd, variance)
high_sd   <- sd(high_density$Density)
high_mean <- mean(high_density$Density)
high_var <-high_sd**2

low_sd   <- sd(low_density$Density)
low_mean <- mean(low_density$Density)
low_var <-low_sd**2
```

```{r}
# Sample size estimation for question
delta <- high_mean - low_mean
n1 <- dim(high_density)[1]
n2 <- dim(low_density)[1]
denom <- sqrt(high_var/n1 + low_var/n2)
Z = delta/denom
n = ((low_var + high_var)*(1.96 + Z))/(delta**2)
```



# Type 1 error probability 
# Power estimation
# Actual Anova Test (Assuming qual variances)
