---
title: "Question 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggpubr)
library(gridExtra)
```

## **Question 2: Do similar data science jobs that involve different programming languages see comparable salaries (i.e. Data Analyst/Data Scientist, Machine Learning Engineer/Machine Learning Scientist)?**

We needed to create groups for the 17 different entries to the Current Job Title descriptions. For this we decided upon grouping titles into 'Scientist' (people probably utilizing higher level math/statistics), 'ResearchAnalyst' (those focused with general data analysis and manipulation), 'Engineer' (titles focused in the data engineering or software), and 'Other' (description unknown).

```{r, Jobtitlegroups}

Scientists <- c("Statistician", "Data Scientist", "Machine Learning Engineer", "Predictive Modeler", "Scientist/Researcher")
ResearchAnalyst <- c("Data Analyst", "Business Analyst", "Data Miner", "Researcher")
Engineer <- c("Software Developer/Software Engineer", "Programmer", "Computer Scientist", "Operations Research Practitioner", "Engineer", "DBA/Database Engineer")

DSdata1 <- DSdata[,] 
DSdata1$JobTitleGroup <- if_else(DSdata1$CurrentJobTitleSelect %in% Scientists, "Scientist", if_else(DSdata1$CurrentJobTitleSelect %in% ResearchAnalyst,'ResearchAnalyst',if_else(DSdata1$CurrentJobTitleSelect %in% Engineer, 'Engineer', 'Other'))) 

```

```{r, setup-dfs}

pythondf <- DSdata1 %>%
  filter(!is.na(CompensationYearUSD), LanguageRecommendationSelect == 'Python')

Rdf <- DSdata1 %>%
  filter(!is.na(CompensationYearUSD), LanguageRecommendationSelect == 'R')

sqldf <- DSdata1 %>%
  filter(!is.na(CompensationYearUSD), LanguageRecommendationSelect == 'SQL')

```

```{r, eda-figures}

rden <- Rdf %>%
  ggplot(aes(x=CompensationYearUSD, fill = JobTitleGroup, color = JobTitleGroup))+ geom_density(alpha = .3) + scale_x_log10() + ggtitle('R') 

qden <-sqldf %>%
  ggplot(aes(x=CompensationYearUSD, fill = JobTitleGroup, color = JobTitleGroup))+ geom_density(alpha = .3) + scale_x_log10() + ggtitle('SQL') 

pden <- pythondf %>%
  ggplot(aes(x=CompensationYearUSD, fill = JobTitleGroup, color = JobTitleGroup))+ geom_density(alpha = .3) + scale_x_log10() + ggtitle("Python")

ggarrange(pden, rden, qden,
          ncol = 1, nrow = 3)
```

From the density chart, we can see that there is a general similarity to the distribution of annual compensation in USD for all the groups within a given language.

### *Group by primary programming language (python, R, SQL..) and test for significant difference among groups through Welch t-test and ANOVA tests. *
