---
title: "Report"
author: "Anmol Srivastava, Juan Solorio, Matthew Rhodes, and Andy De La Fuente"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::read_chunk("../Code/01-Question1.Rmd")

```

# Title NUMBER UR PAGES PDF 

## Abstract


## Introduction
aims + questions, background material/research briefly 
For our project we decided to go with the theme of Data Science and Pay. To help us answer this question we found a 
dataset from Kaggle listed here: https://www.kaggle.com/ikleiman/data-scientists-salaries-around-the-world/data.
See the next section for our description of the data.
After analysis of the data during the initial exploration, we noticed that the
data should be split into two initial groups (people who are working from commission and those on salary).
This split was made because the data contained a large amount of answers that were less than an assumed salary 
would be (40k+). There were also data points that were obvious `outliers` (people who earn 1,000,000 or more) the
maximum salary was 28,000,000,000 so
we decided to cap those values because we conluded that nobody is making more than 500,000 salary. Another thing
that should be noted is that we are only considering people that are employed.
## Data Set Description
source, collection methods, study design (randomized exp, obs, survey, etc)




## Statistical Methods
methods, discussion of assumptions, explanation for why methods appropriate 
Anova, Welch T-test, linear regression

## Results 
graphs, tables, descriptive info about data and results of analysis 
# Results from Question 1
```{r, echo =FALSE}
invisible(knitr::purl("../Code/01-Question1.Rmd", output="temp", quiet=TRUE))
knitr::read_chunk("temp")
```

```{r ref.label='Q1_data'}
```

```{r ref.label='testing'}
```

```{r ref.label='pre_plots'}
```

```{r ref.label='Question_1_plots'}
```

```{r echo=FALSE}
unlink("temp")
```

## Discussion
summarize results and conclusions 
limits of analyses 
limits of data too 

## References
brief 

## Appendices 
more technical aspects of analyses, any other tidbits 
